---
title: "Data Due Diligence"
author: "Cameron Gallien"
date: "`r Sys.Date()`"
output: pdf_document
---
Importing data and summarizing.

```{r}
library(tidyverse)
library(dplyr)
CustomerDF = read_csv("Customer_Dataset_Data.csv")
summary(CustomerDF)
```
#HHIncome -> Numeric

```{r}
CustomerDF <-CustomerDF %>% mutate(HHIncome_Dollars = parse_number(HHIncome))
print(head(CustomerDF$HHIncome_Dollars))
print(summary(CustomerDF$HHIncome_Dollars))
```
#Cleaning CarOwnership & CarBrand

<!-- CustomerDF$CarOwnership_Clean <- CustomerDF$CarOwnership -->
<!-- CustomerDF$CarOwnership_Clean <- na_if(CustomerDF$CarOwnership_Clean, -1) -->
<!-- CustomerDF$CarBrand_Clean <- CustomerDF$CarBrand -->
<!-- CustomerDF$CarBrand_Clean <- na_if(CustomerDF$CarBrand_Clean, -1) -->

#CarValue -> Numeric, take out -1000s

```{r}
CustomerDF$CarValue_Dollars <- CustomerDF$CarValue
CustomerDF['CarValue_Dollars'][CustomerDF['CarValue_Dollars']=='($1,000.00)'] <- '0'

CustomerDF <-CustomerDF %>% mutate(CarValue_Dollars = parse_number(CarValue_Dollars))
print(head(CustomerDF$CarValue_Dollars))
print(summary(CustomerDF$CarValue_Dollars))
```

#CardSpendMonth -> Numeric, NA's to 0

```{r}
CustomerDF <-CustomerDF %>% mutate(CardMonth_Dollars = parse_number(CardSpendMonth))
print(head(CustomerDF$CardMonth_Dollars))
print(summary(CustomerDF$CardMonth_Dollars))
```

```{r}
CustomerDF = CustomerDF %>% mutate(CardMonth_Dollars = replace_na(CustomerDF$CardMonth_Dollars,replace=0))
print(summary(CustomerDF$CardMonth_Dollars))
```

#VoiceLastMonth -> Numeric

```{r}
CustomerDF <-CustomerDF %>% mutate(VoiceMonth_Dollars = parse_number(VoiceLastMonth))
print(head(CustomerDF$VoiceMonth_Dollars))
print(summary(CustomerDF$VoiceMonth_Dollars))
```

#VoiceOverTenure -> Numeric

Going to take out NA values later when conducting analyses

```{r}
CustomerDF <-CustomerDF %>% mutate(VoiceTenure_Dollars = parse_number(VoiceOverTenure))
print(head(CustomerDF$VoiceTenure_Dollars))
print(summary(CustomerDF$VoiceTenure_Dollars))
```

#EquipmentLastMonth -> Numeric, NAs -> 0

```{r}
CustomerDF <-CustomerDF %>% mutate(EquipMonth_Dollars = parse_number(EquipmentLastMonth))
print(head(CustomerDF$EquipMonth_Dollars))
print(summary(CustomerDF$EquipMonth_Dollars))
```
```{r}
CustomerDF = CustomerDF %>% mutate(EquipMonth_Dollars = replace_na(CustomerDF$EquipMonth_Dollars,replace=0))
print(summary(CustomerDF$EquipMonth_Dollars))
```

#EquipmentOverTenure -> Numeric, NAs to 0

```{r}
CustomerDF <-CustomerDF %>% mutate(EquipTenure_Dollars = parse_number(EquipmentOverTenure))
print(head(CustomerDF$EquipTenure_Dollars))
print(summary(CustomerDF$EquipTenure_Dollars))
```
```{r}
CustomerDF = CustomerDF %>% mutate(EquipTenure_Dollars = replace_na(CustomerDF$EquipTenure_Dollars,replace=0))
print(summary(CustomerDF$EquipTenure_Dollars))
```

#DataLastMonth -> Numeric, NAs -> 0

```{r}
CustomerDF <-CustomerDF %>% mutate(DataMonth_Dollars = parse_number(DataLastMonth))
print(head(CustomerDF$DataMonth_Dollars))
print(summary(CustomerDF$DataMonth_Dollars))
```

```{r}
CustomerDF = CustomerDF %>% mutate(DataMonth_Dollars = replace_na(CustomerDF$DataMonth_Dollars,replace=0))
print(summary(CustomerDF$DataMonth_Dollars))
```

#DataOverTenure-> Numeric, NAs -> 0

```{r}
CustomerDF <-CustomerDF %>% mutate(DataTenure_Dollars = parse_number(DataOverTenure))
print(head(CustomerDF$DataTenure_Dollars))
print(summary(CustomerDF$DataTenure_Dollars))
```

```{r}
CustomerDF = CustomerDF %>% mutate(DataTenure_Dollars = replace_na(CustomerDF$DataTenure_Dollars,replace=0))
print(summary(CustomerDF$DataTenure_Dollars))
```

<!-- 
<!-- CustomerDF$Internet_Cleaned <-CustomerDF$Internet -->
<!-- CustomerDF$Internet_Cleaned <- replace(CustomerDF$Internet_Cleaned, CustomerDF$Internet_Cleaned=="2", NA) -->
<!-- CustomerDF$Internet_Cleaned <- replace(CustomerDF$Internet_Cleaned, CustomerDF$Internet_Cleaned=="3", NA) -->
<!-- CustomerDF$Internet_Cleaned <- replace(CustomerDF$Internet_Cleaned, CustomerDF$Internet_Cleaned=="4", NA) -->

#Create new column - All Services

```{r}
CustomerDF$AllServices <- ifelse(CustomerDF$DataTenure_Dollars>0 & CustomerDF$VoiceTenure_Dollars>0 & CustomerDF$EquipTenure_Dollars>0,TRUE, FALSE)
```

# Create new column - All Services total

```{r}
CustomerDF <- CustomerDF %>% rowwise() %>% mutate(AllSvcTotal = sum(c(DataTenure_Dollars, VoiceTenure_Dollars, EquipTenure_Dollars)))

```
# Create new column - All Services avg month

```{r}
CustomerDF <- CustomerDF %>% rowwise() %>% mutate(AllSvcMonth = sum(c(DataTenure_Dollars, VoiceTenure_Dollars, EquipTenure_Dollars)/PhoneCoTenure))

```
# New column - HHIncome/HHSize
```{r}
CustomerDF$IncomePerMember <- CustomerDF$HHIncome_Dollars/CustomerDF$HouseholdSize
```
# Total Tech New Column
```{r}
df1 <- data.frame(CustomerDF$CustomerID, CustomerDF$OwnsPC, CustomerDF$OwnsMobileDevice, CustomerDF$OwnsGameSystem, CustomerDF$OwnsFax)
df1$totaltech <- apply(df1, 1, function(x) length(which(x=="Yes")))

CustomerDF <- data.frame(CustomerDF, df1$totaltech)
names(CustomerDF)[74] <- "TotalTech"
```


# Visualizations

```{r}
library(ggplot2)
```

```{r}
RegionHist <- ggplot(data=CustomerDF) + geom_bar(aes(x=Region))
print(RegionHist)
```
```{r}
JobHist <- ggplot(data=CustomerDF) + geom_bar(aes(x=JobCategory))
print(JobHist)
```

```{r}
TotalTechHist <- ggplot(data=CustomerDF) + geom_bar(aes(x=TotalTech)) +
  labs(x='Total Additional Pieces of Technology')
print(TotalTechHist)
```

```{r}
IncomeByJobCategory <- ggplot(data=CustomerDF) + geom_boxplot(aes(x = as.factor(JobCategory),y=AllSvcMonth)) + 
  labs(x='Job Category',
       y='Average Spent per Month in Dollars')
print(IncomeByJobCategory)

```

```{r}
AverageSpentByRegion <- ggplot(data=CustomerDF) + geom_boxplot(aes(x = as.factor(Region),y=AllSvcMonth)) +
  labs(x='Region',
       y='Average Spent per Month in Dollars')
print(AverageSpentByRegion)

```

```{r}
Education_Spend <- ggplot(data=CustomerDF) +geom_point(aes(x=EducationYears,y=AllSvcMonth
)) +
  labs(x='Years of Education',
       y='Average Spent per Month in Dollars')
print(Education_Spend)
```

```{r}
AmountSpendByTotalTech <- ggplot(data=CustomerDF) + geom_boxplot(aes(x = as.factor(TotalTech
),y=AllSvcMonth)) +
  labs(x='Total Additional Pieces of Technology',
       y='Average Spent per Month in Dollars')
print(AmountSpendByTotalTech)
```

```{r}
AmountSpendByServices <- ggplot(data=CustomerDF) + geom_boxplot(aes(x = as.factor(AllServices
),y=AllSvcMonth)) +
  labs(x='Subscribes to All Services',
       y='Average Spent per Month in Dollars')
print(AmountSpendByServices)
```
```{r}
HHIncome_Spend <- ggplot(data=CustomerDF) +geom_point(aes(x=HHIncome_Dollars,y=AllSvcMonth)) +
coord_cartesian(xlim = c(0, 1000000)) +
  labs(x='HHIncome in Dollars',
       y='Average Spent per Month in Dollars',
  )
print(HHIncome_Spend)
```

```{r}
Debt_Spend <- ggplot(data=CustomerDF) +geom_point(aes(x=DebtToIncomeRatio,y=AllSvcMonth)) +
  labs(x='Debt To Income Ratio',
       y='Average Spent per Month in Dollars')

print(Debt_Spend)
```



