---
title: "Week 7 - Rules-based Clustering for Segmentation"
author: "Chris Healey"
output:
  html_document:
    toc: true
    theme: united
---

I want to use a common ML dataset, Titanic passengers, for our examples this week.

Our main goal is to identify characteristics of passenger survival using rules-based clustering (covering continuous and categorical features).


# Creating rules
In this section, we try to identify clusters with a few simple rules.  

## Load data

```{r}
library(tidyverse)
library(ggplot2)
```
```{r}
#load data
DDD <- read_csv("CustomerDF_Clean.csv")
```

## Use cut to create groups 
```{r}
PCT <- ggplot(data=DDD) + geom_histogram(aes(x=PhoneCoTenure))
print(PCT)
```
Let's create a rule based on Fare using **cut**, turning the continuous variable Fare into a discrete one called Group

```{r}
DDD$PCTGroup <- cut(DDD$PhoneCoTenure, breaks=c(0,24,48,72), labels=c("LowTen","MedTen","HighTen"))
DDD$TTGroup <- cut(DDD$TotalTech, breaks=c(-1,2,4), labels=c("LowTech", "HighTech"))
DDD$HIGroup <- cut(DDD$HHIncome_Dollars, breaks=c(0,38000,67000, 1073000), labels=c("LowHI","MedHI","HighHI"))
DDD$MonthSvcGroup <- cut(DDD$AllSvcMonth, breaks=c(0,51, 300), labels=c("LowMth","HighMth"))
#DDD$AgeGroup <- cut(DDD$Age, breaks=c(0,47, 80), labels=c("Young","Old"))
DDD$AgeGroup <- cut(DDD$Age, breaks=c(0,39,59, 80), labels=c("Young","Mid", "Old"))
DDD$TVGroup <- cut(DDD$TVWatchingHours, breaks=c(0,20, 37), labels=c("High","Low"))
DDD$HHSizeGroup <- cut(DDD$HouseholdSize, breaks=c(0,2,5, 10), labels=c("LowSize","MedSize","HighSize"))
DDD$PetGroup <- cut(DDD$NumberPets, breaks=c(-1,0,2, 10), labels=c("None","Few","Many"))

```

We can combine this rule with the Sex variable to create 6 segments

```{r}
#IncomeVSAge
DDDGrps <- DDD %>% mutate(segment = case_when(
      HIGroup == "LowHI" & AgeGroup == "Young" ~ 1,
      HIGroup == "MedHI" & AgeGroup == "Young" ~ 2,
      HIGroup == "HighHI" & AgeGroup == "Young" ~ 3,
      HIGroup == "LowHI" & AgeGroup != "Young" ~ 4,
      HIGroup == "MedHI" & AgeGroup != "Young" ~ 5,
      TRUE ~ 6
    ))

#df %>% write_csv("DDDGrps.csv")
```

```{r}
#MonthlyVSAge
DDDGrps <- DDD %>% mutate(segment = case_when(
      MonthSvcGroup == "LowMth" & AgeGroup == "Young" ~ 1,
      MonthSvcGroup == "HighMth" & AgeGroup == "Young" ~ 2,
      MonthSvcGroup == "LowMth" & AgeGroup != "Young" ~ 3,
      TRUE ~ 4
    ))

#df %>% write_csv("output_of_segmentation.csv")
```

```{r}
#MonthlyVSAge
DDDGrps <- DDD %>% mutate(segment = case_when(
      NewsSubscriber == "No" & AgeGroup == "Young" ~ 1,
      NewsSubscriber == "Yes" & AgeGroup == "Young" ~ 2,
      NewsSubscriber == "No" & AgeGroup != "Young" ~ 3,
      TRUE ~ 4
    ))

#df %>% write_csv("output_of_segmentation.csv")
```

```{r}
#NewsVSAge
DDDGrps <- DDD %>% mutate(segment = case_when(
      AgeGroup == "Young" & NewsSubscriber == "No" ~ 1,
      AgeGroup == "Mid" & NewsSubscriber == "No" ~ 2,
      AgeGroup == "Old" & NewsSubscriber == "No" ~ 3,
      AgeGroup == "Young" & NewsSubscriber != "No" ~ 4,
      AgeGroup == "Mid" & NewsSubscriber != "No" ~ 5,
      TRUE ~ 6
    ))

#df %>% write_csv("output_of_segmentation.csv")
```

```{r}
#MaritalVSAge
DDDGrps <- DDD %>% mutate(segment = case_when(
      MaritalStatus == "Married" & AgeGroup == "Young" ~ 1,
      MaritalStatus == "Unmarried" & AgeGroup == "Young" ~ 2,
      MaritalStatus == "Married" & AgeGroup != "Young" ~ 3,
      TRUE ~ 4
    ))

#df %>% write_csv("output_of_segmentation.csv")
```


## Segment profiles
For the segment profiles, I summarize the results of each segment (using the original dataset columns -- without the scaling).
```{r}
DDD %>% add_count(AgeGroup) %>%
  group_by(AgeGroup,n) %>% 
  summarise_all("mean")
```

```{r}
DDDGrps %>% select(-c("NewsSubscriber", "AgeGroup")) %>% add_count(segment) %>%
  group_by(segment,n) %>% 
  summarise_all("mean")
```

## Example 2-D Visualization
Here's visualization of the segments featuring two real-valued variables.
```{r}
g <- ggplot(DDDGrps)+geom_point(aes(x=PhoneCoTenure,y=Age,colour=as.factor(segment)))
print(g)
```

```{r}
PCT <- ggplot(data=DDDGrps) + geom_histogram(aes(x=segment))
print(PCT)
```